---
title: "B5-Global_Change_Impact"
author: "Ridwan Shittu"
date: "2023-06-16"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: 
      collapsed: false
      smooth_scroll: false 
    
    

---



# Species distribution modeling of Culex pipiens in Germany
## Objectives
Model the bioclimatic niche of a species endemic to Europe and
create a set of projections under a collection of climate change scenarios.
A further objective is to measure the sensitivity of the species to climate
change under the different scenarios.

## Data
This example relies on presence-only data downloaded from GBIF,
which will then require the creation of a set, or several sets, of pseudoabsence
data. The explanatory variables are raster grid data downloaded
from the WorldClim datacenter.

### Modeling Steps
- Loading and formatting the presence- only data
- Loading and formatting the raster data
- Building a range of models and ensemble models using biomod2
- Decomposing the models’ variability (predictive ability / predictions)
- Projections under current and future conditions
- Species’ range change estimates.


```{r, message=FALSE, warning=FALSE}

library(rgbif)
library(biomod2)
library(ggplot2)
library(gridExtra)

```


## Obtaining Species Data from a Datacenter
```{r specie data}
culex_sp <- name_suggest(q='Culex pipiens', rank = 'species', limit=10000)
culex_sp

#(culex_sp <- culex_sp[grepl("^Culex pipiens", culex_sp$data$canonicalName), ])

# find the taxonKey(usageKey)
name_backbone('Culex pipiens')$usageKey 
occ_search(scientificName = "Culex pipiens")



```
```{r}
data <- occ_search(taxonKey = 1652991, country="DE", fields = "minimal")
# data <- occ_search(taxonKey = 1652991, country="DE", fields = "minimal", limit=2000)

#data$data$scientificName <- sub(" ", ".", data$data$scientificName ) # not 
#(culex_occ <- unique(data$data$scientificName))

#total number of occurrence
sort(table(data$data$scientificName), decreasing = T)

```
## Environmental Data 
```{r}
# get WorldClim environmental variables

dir.create("WorldClim_data", showWarnings = F)
library(geodata)
?geodata


WC20<- worldclim_global(var="bioc", res = 10, path=paste0(getwd(),"/WorldClim_data/"))

#load the future datasets # explain the features 
GCM70 <-cmip6_world("MPI-ESM1-2-LR", "585", "2061-2080", var="bioc", res=10, path=paste0(getwd(),"/WorldClim_data/"))

GCM50 <-cmip6_world("MPI-ESM1-2-LR", "585", "2041-2060", var="bioc", res=10, path=paste0(getwd(),"/WorldClim_data/")) 


list.files("WorldClim_data/")
list.files("WorldClim_data/wc2.1_10m/")

```
## Environmental variable selection

```{r}
head(data$data)
culex_occ=data$data
head(culex_occ)
colnames(culex_occ)
unique(culex_occ$scientificName)
culex_occ <- culex_occ[culex_occ$scientificName=="Culex pipiens Linnaeus, 1758", ]

dim(culex_occ)


library(gtools)

#stack the environmental variables
bioclim_world <-list.files("WorldClim_data/wc2.1_10m/", patter="bio_", full.names = T)
bioclim_world <- mixedsort(bioclim_world )
bioclim_world <- raster::stack(bioclim_world)


# download germany boundaries
DE <- gadm(country="Germany", level=0, resolution = 1, version="4.1",  path=paste0(getwd(),"/WorldClim_data/"))
plot(DE)
class(DE)
class(bioclim_world)
bioclim_DE <- raster::crop(bioclim_world, DE)
```

